Impact of Satellite Constellations on the Legacy Survey of Space and Time

### 2022-10-05 ms.tex

However, one challenge for the LSST is that increasing numbers of bright
Low-Earth-Orbit (LEO) satellites (e.g., Starlink) are being launched, which could leave streaks
in astronomical pointings. Over the last three years, many astronomers have raised concerns
about the impact of this on the LEO ecosystem and astronomical surveys \citep{lawrence22,tyson20}. LEO satellites
are visible from Earth because they reflect sunlight, especially during twilight. As the
illuminated LEO satellites move across the field of view of an astronomical pointing, they leave
a streak in the image, which negatively impacts the scientific value of the pointing. Astronomers
predict there will be at least one satellite streaks in the majority of exposures of the Rubin
Observatory \citep{lawrence22}. % XXX-but, our results say OneWeb streaks 10% of Rubin images. We shouldn't repeat claims we know are wrong. And this is just citing tyson20, which does not say the majority of exposures would have a streak. 
An example of satellite streaks in an astronomical image is shown in Figure \ref{fig-streaks}.

@@

\begin{figure}[ht!]
%\includegraphics{} 
\caption{Example image of a satellite streak \label{fig-streaks}}
\end{figure}

There have been efforts to reduce the impact of satellite streaks in astronomical pointings.
Satellite companies like SpaceX have worked on darkening the exterior of satellites so they will
be less illuminated [CITE SPACEX BLOG POST]. However, even with extra darkening, the Starlink satellites are still one
magnitude too bright for the observatory cameras, meaning these darkened satellite movements
will still be captured by the camera. Astronomers have worked on algorithms for masking
satellite trails in images, but covering the outer rim of the trails without losing extra pixels
remains a challenge \citep{tyson20}.
The rapid increase in population of LEO satellites threatens to compromise the quality
and scientific value of the LSST images and also necessitates expensive computational power to
mask the trails. Thus, we explore another option: incorporating known commercial satellites into
the Rubin Scheduler so the worst of them may be avoided. In this project, we create realistic
simulated forecasts of satellite trajectories and brightness, build a tool that uses that data to create
new scheduler constraints, and test the impact of the new scheduler on various Rubin LSST
observing programs.

%\section{Constellations}

\section{Methodology}
We begin by creating realistic forecasts of three different commercial satellite constellations: StarlinkV1 (4408 satellites, altitude 540-570 km), OneWeb (6372 satellites, altitude 1200 km), and StarlinkV2 (29988 satellites, altitude 340-614 km). All three of these constellations are highly likely to launch as designed (XXX-lol, no), or close to it, as both operators presently have functional satellites in orbit and are working to build and launch more. These simulated satellite
constellations are illustrated in Figure \ref{fig-simulated-constellations}.

\begin{figure}[ht!]
\epsscale{0.35}
\plotone{plots/sats_slv1.pdf}
\plotone{plots/sats_slv2.pdf}
\plotone{plots/sats_ow.pdf}\\
\plotone{plots/sats_slv1_late.pdf}
\plotone{plots/sats_slv2_late.pdf}
\plotone{plots/sats_ow_late.pdf}
\epsscale{1}
\caption{Hammer projection of the altitude and azimuth positions of our simulated satellite constellations as seen from Rubin Observatory on October 1, 2023 (sun altitude -18 degrees). Red points show un-illuminated satellites, blue points show those illuminated by the sun, and black points are illuminated and above the Rubin altitude limit. \label{fig-simulated-constellations} XXX--could add a second row to show how nothing is up when the sun is very low.}
\end{figure}

@@

\begin{figure}[ht!]
\epsscale{0.3}
\plotone{plots/starlink_tles_v1_0.00_basisfunc.pdf}
\plotone{plots/starlink_tles_v2_0.00_basisfunc.pdf}
\plotone{plots/oneweb_tles_0.00_basisfunc.pdf}
\epsscale{1}
\caption{Example of satellite avoidance maps constructed for the scheduler. The maps have been rotated so zenith is in the center of the image. By varying the weight placed on these maps, the scheduler will more actively avoid regions of the sky where satellites could streak images. \label{fig-simulated-scheduler}}
\end{figure}

@@

XXX--Should we add in the results from looking at a run with twilight NEO survey observations? Doesn't flow well with everything else, but I think it is a result that will help some folks calm down a bit that we aren't going to miss tons of killer asteroids even with large constellations.

@@

\begin{figure}[ht!]
\epsscale{0.35}
\plotone{plots/slv1coadd.pdf}
\plotone{plots/slv2coadd.pdf}
\plotone{plots/onewcoadd.pdf}
\epsscale{1}
\caption{The trade off between pixels lost to streaks and final median coadded depth of the survey. Points to the right of the dashed line show improvement while those to the left signify the regained pixels do not adequatly compensate for the lost depth. In no situation are we able to prevent depth loss in at least some filters. \label{fig-trade-off}}
\end{figure}

@@

@@

>>>>

However, one challenge for the LSST is that increasing numbers of bright Low-Earth-Orbit (LEO) satellites (e.g., Starlink) are being launched, which could leave streaks in astronomical pointings. Over the last three years, many astronomers have raised concerns
about the impact of this on the LEO ecosystem and astronomical surveys \citep{lawrence22,tyson20}. LEO satellites are visible from Earth because they reflect sunlight, especially during twilight. As the illuminated LEO satellites move across the field of view of an astronomical pointing, they leave a streak in the image, which negatively impacts the scientific value of the pointing.

@@

There have been efforts to reduce the impact of satellite streaks in astronomical pointings. Satellite companies like SpaceX have worked on darkening the exterior of satellites so they will be less illuminated\footnote{\url{https://api.starlink.com/public-files/BrightnessMitigationBestPracticesSatelliteOperators.pdf}}. However, even with extra darkening, the Starlink satellites are still one magnitude too bright for the observatory cameras, meaning these darkened satellite movements will still be captured by the camera. Astronomers have worked on algorithms for masking satellite trails in images, but covering the outer rim of the trails without losing extra pixels remains a challenge \citep{tyson20}. The rapid increase in population of LEO satellites threatens to compromise the quality and scientific value of the LSST images and also necessitates expensive computational power to mask the trails. Thus, we explore another option: incorporating known commercial satellites into the Rubin Scheduler so the worst of them may be avoided. In this project, we create realistic simulated forecasts of satellite trajectories and brightness, build a tool that uses that data to create new scheduler constraints, and test the impact of the new scheduler on various Rubin LSST observing programs.

\section{Methodology}
We begin by creating realistic forecasts of three different commercial satellite constellations: StarlinkV1 (4,408 satellites, altitude 540-570 km), OneWeb (6,372 satellites, altitude 1200 km), and StarlinkV2 (29,988 satellites, altitude 340-614 km). All three of these constellations are highly likely to launch as designed (XXX-lol, no), or close to it, as both operators presently have functional satellites in orbit and are working to build and launch more. These simulated satellite constellations are illustrated in Figure \ref{fig-simulated-constellations}.

% TODO: add three more panels on the top that show each simulated constellation
% as it appeared in the poster
% TODO: make label fonts on each figure GINORMOUS so they are legible when shrunk,
% and delete any extraneous labels
\begin{figure}[ht!]
\epsscale{0.35}
\plotone{plots/starlinkv1.pdf}
\plotone{plots/starlinkv2.pdf}
\plotone{plots/oneweb.pdf}
\plotone{plots/sats_slv1.pdf}
\plotone{plots/sats_slv2.pdf}
\plotone{plots/sats_ow.pdf}\\
\plotone{plots/sats_slv1_late.pdf}
\plotone{plots/sats_slv2_late.pdf}
\plotone{plots/sats_ow_late.pdf}
\epsscale{1}
\caption{Hammer projection of the altitude and azimuth positions of our simulated satellite constellations as seen from Rubin Observatory on October 1, 2023 (sun altitude -18 degrees). Red points show un-illuminated satellites, blue points show those illuminated by the sun, and black points are illuminated and above the Rubin altitude limit. The second row is the same, but for 6 hours later when the sun has reached and altitude of -50 degrees. Because the Starlink constellations are lower altitude, they have no illuminated satellites in the middle of the night, while the OneWeb constellation has only one illuminated satellite above the Rubin altitude limit. \label{fig-simulated-constellations} }
\end{figure}

@@

\begin{figure}[ht!]
\epsscale{0.37}
\plotone{plots/starlink_tles_v1_0.00_basisfunc.pdf}
\plotone{plots/starlink_tles_v2_0.00_basisfunc.pdf}
\plotone{plots/oneweb_tles_0.00_basisfunc.pdf}
\epsscale{1}
\caption{Example of satellite avoidance maps constructed for the scheduler. The maps have been rotated so zenith is in the center of the image. By varying the weight placed on these maps, the scheduler will more actively avoid regions of the sky where satellites could streak images. \label{fig-simulated-scheduler}}
\end{figure}

@@

@@

\begin{figure}[ht!]
\epsscale{0.7}
%\plotone{plots/slv1coadd.pdf}
%\plotone{plots/slv2coadd.pdf}
%\plotone{plots/onewcoadd.pdf}\\
\plotone{plots/streak_length_v_deltam.pdf}
\epsscale{1}
\caption{The trade off between pixels lost to streaks and final median coadded depth of the survey. Points to the right of the dashed line show improvement while those to the left signify the regained pixels do not adequatly compensate for the lost depth. In no situation are we able to prevent depth loss in at least some filters. \label{fig-trade-off}
XXX--the top panels have more info, but I kinda like the lower for the simplicity. Maybe just the lower and say all the filters follow a similar trend? Or a plot for g and a plot for i?
xxx--MR deleted the top panels, and thinks we should avoid discussing filters/bands too much because people will ask whether we are simulating sat brightness as a function of wavelength... which we definitely aren't}
\end{figure}

@@

xxx--Add short statement about impacts on near-Earth object (NEO) searches, and also a figure that Peter will make.
XXX--Add in the results from looking at a run with twilight NEO survey observations. I think it is a result that will help some folks calm down a bit that we aren't going to miss tons of killer asteroids even with large constellations.

@@

xxx--Consider saying something like only 10\% of images will have a streak from these three constellations, whereas Lawrence et al. 2022 says ``a majority''. Some of this is probably based on old Starlink info with higher altitudes.
